Do expert pianists adapt their pedagogical demonstrations to novices' skills?

### 2022-12-01 paper/manuscript.Rmd
reflected some feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002]. Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007].

Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners. @fukuyama_2015 revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation. For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions. The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.

In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics. For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had the intention to teach dynamics. Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance. If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.

In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners. Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control. The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review). All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics. In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (*Fig \ref{fig:stim}*). We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings. Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.

@@

The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices. We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music. The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills. On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills. The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.

We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students. Even compared with the tempo of the baseline performance, they did not change the tempo either. These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics. Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).

For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings. This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings. However, we did not find significant results in terms of dynamics. Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.

Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation. One possibility would be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation. In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music. Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it. Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.

One of the limitations of the current study is that participants did not have an idea about what the ideal performance was. As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors. One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.

The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo. This was to make sure that we manipulated articulation and dynamics features of the recordings only. However, all the recordings might not have sounded as if unskillful students were playing. In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.

>>>>

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions [@pezzulo_2019], including teaching contexts. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths [@brand_2002]. Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners [@mcellin_2017; @uther_2007]. For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own [@mcellin_2017].

Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners. @fukuyama_2015 found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills. For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions. To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.

To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners. In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics. In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics. When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato. In that study, teachers were instructed to demonstrate techniques to hypothetical learners and they did not hear novices' particular performances. This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way caregivers have been shown to adapt demonstrations of actions to their infants' skills. If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.

To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review). All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics. According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (*Fig \ref{fig:stim}*). We generated four types of recordings. In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented. This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.

@@

The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novice whom they are trying to teach. We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music. The recordings where the two techniques were implemented were supposed to resemble the performances of students with good skills. The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics. The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.

Expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings. This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings. However, we did not find significant results in terms of dynamics. Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.

We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students. Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context. These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics. Although slower demonstration is generally considered to be used for teaching [@mcellin_2017; @saint-georges_2013; @schaik_2019], specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.

Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation. One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation. In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music. Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it. Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.

One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind. As some participants reported that "(students had) very different interpretations of the melody" or "all kids played rhythmically correct" in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors. One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.

The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo. This was to make sure that only articulation and dynamics features of varied. However, some of the recordings might not have sounded as if unskillful students were playing. In order to create more realistic novice performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask beginner pianists to play the piece and extract the temporal features of their performances.

### 2022-12-01 paper/manuscript.tex
reflected some feedback
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002). Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007).

Furthermore, recent evidence has shown that such pedagogical modulations were sensitive to the skill levels of learners. Fukuyama et al. (2015) revealed mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation. For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have enough motor skills to produce such actions. The results indicated that adults monitored infants' actions and abilities, and adapted their demonstration accordingly.

In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics. For example, expert pianists exaggerated dynamics contrast (i.e., making larger contrast between forte and piano) when they had the intention to teach dynamics. Such exaggeration is considered to be most effective for those who fail to produce the techniques in their performance. If teaching is an adaptive and scaffolding process, it can be expected that experts should not use such exaggeration if novices became able to produce them.

In the current study, we investigated whether and how expert pianists adapted their performance depending on the skill levels of novice learners. Instead of recruiting actual novice pianists, we created artificial recordings to manipulate the skill levels of novices to maximise the experimental control. The recordings were made based on the performance data from our previous experiment (Tominaga et al., under review). All the recordings had the same musical structure (i.e., the same piece) but had different implementations of articulation and dynamics. In the ideal performance, both articulation and dynamics should be implemented as notated on the sheet music (\emph{Fig \ref{fig:stim}}). We made the recordings so that both techniques were implemented in the first quarter, only articulation was implemented in the second quarter, only dynamics was implemented in the third quarter and neither of them was implemented in the last quarter of the recordings. Therefore, in this experiment, the skill levels of novices corresponded to the implementation of articulation and dynamics.

@@

The present study investigated whether and how expert pianists adapt their performance depending on the skill levels of novices. We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music. The recordings where the two techniques were implemented were supposed to replicate the performances of students with high skills. On the other hand, the recordings where none of the two techniques was implemented were supposed to replicate the performances of students with low skills. The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.

We found that expert pianists did not modulate their tempo (IOIs) depending on the skill levels of students. Even compared with the tempo of the baseline performance, they did not change the tempo either. These findings indicated that tempo was not employed specifically for teaching because tempo itself was relevant to neither articulation nor dynamics. Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing tempo might give another interpretation of music (e.g., expressive timing).

For KOT, expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings. This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings. However, we did not find significant results in terms of dynamics. Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.

Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation. One possibility would be that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation. In our previous experiment, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music. Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to teach it. Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.

One of the limitations of the current study is that participants did not have an idea about what the ideal performance was. As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors. One solution would be to create a model performance where both articulation and dynamics was implemented and ask participants (expert pianists) to listen to it before starting the current experiment so that they can detect errors by comparing each recording with the model performance.

The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo. This was to make sure that we manipulated articulation and dynamics features of the recordings only. However, all the recordings might not have sounded as if unskillful students were playing. In order to create more realistic unskillful performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask unskillful students to play the piece and extract the temporal feature of their performances.

>>>>

Exaggeration of movement is often used to signal a communicative intent to others during real-time interactions (Pezzulo et al., 2019), including teaching contexts. For instance, when caregivers were teaching how to use a novel toy to their infants, they tended to exaggerate the movement paths (Brand, Baldwin, \& Ashburn, 2002). Such exaggeration was observed not only towards infant learners but also when adults were interacting with adult learners (McEllin, Knoblich, \& Sebanz, 2017; Uther, Knoll, \& Burnham, 2007). For example, participants demonstrating to a learner how to play a particular melody on a xylophone increased the amplitude of their movements, compared to playing the same melody on their own (McEllin et al., 2017).

Recent evidence from parent-infant interactions suggests that such pedagogical action modulations are sensitive to the skill levels of learners. Fukuyama et al. (2015) found that mothers who were demonstrating how to use nesting cups to their infants dynamically modulated their demonstration according to the infants' object manipulation skills. For example, the mothers exaggerated their movements when their infants failed to produce expected actions whereas the mothers did not produce such exaggeration when their infants did not have the motor skills needed to produce the required actions. To the best of our knowledge, it is unknown whether people teaching skills to adult learners also adjust their demonstrations as a function of the learners' demonstrated skills.

To address this question, in the current study we investigated whether and how expert pianists adapt their pedagogical demonstrations for learners depending on the skill level of the novice learners. In our previous research (Tominaga et al., under review), we demonstrated that expert pianists modulated their performance when they were asked to teach musical expressive techniques such as articulation and dynamics. In particular, expert pianists exaggerated the contrast between different dynamics (i.e., they made a larger contrast between forte and piano) when they had the intention to demonstrate dynamics. When they had the intention to demonstrate articulation, they exaggerated articulation by producing shorter staccato. In that study, teachers were instructed to demonstrate techniques to hypothetical learners and they did not hear novices' particular performances. This leaves open the question whether pianists would make specific adjustments depending on learners' skills in the way caregivers have been shown to adapt demonstrations of actions to their infants' skills. If teachers modulate their demonstrations based on learners' skills, their exaggeration of particular aspects should depend on the learners' actual performance with regard to the techniques to be acquired.

To have experimental control over learners' demonstrated proficiency in implementing particular techniques, we generated artificial recordings of novice piano performances instead of recruiting actual novice pianists The recordings were made based on the performance data from our previous study (Tominaga et al., under review). All the recordings were of the same same piece but the recordings differed in the implementation of articulation and dynamics. According to the notated expressions in the sheet music, proper performance of the piece required specific modulations with respect to both articulation and dynamics (\emph{Fig \ref{fig:stim}}). We generated four types of recordings. In one quarter of the recordings, both techniques were implemented; in one quarter, neither dynamics nor articulation were implemented; in the two remaining quarters, the piece was played with either only articulation or only dynamics implemented. This allowed us to investigate whether expert pianists in the role of teachers would make specific adjustments in response to the learners' demonstrated skills.

@@

The present study investigated whether and how expert pianists adapt their performance depending on the displayed skills of novice whom they are trying to teach. We created artificial recordings to manipulate the skill levels by modifying the implementation of two musical expressive techniques (i.e., articulation and dynamics) notated on the sheet music. The recordings where the two techniques were implemented were supposed to resemble the performances of students with good skills. The recordings where neither of the two techniques was implemented were supposed to resemble the performances of students with lower skills who had not yet learnt how to implement articulation and dynamics. The recordings where either of the two techniques was implemented but the other was missing were equivalent to the performances of students with intermediate skills.

Expert pianists exaggerated legato and staccato when articulation was not implemented in the recordings. This is in line with our predictions that experts would exaggerate only the relevant aspects of the performance if a particular technique was missing in the recordings. However, we did not find significant results in terms of dynamics. Instead, we found that expert pianists produced softer piano when articulation was implemented in the recordings, indicating that participants exaggerated piano only when there was no problem with regard to articulation in the recordings.

We also found that expert pianists did not modulate their tempo (IOIs) depending on the displayed skill levels of students. Also compared with the tempo of the baseline performance, they did not change the tempo in the teaching context. These findings indicated that tempo was not employed specifically teaching, likely because tempo itself was not relevant to neither articulation nor dynamics. Although slower demonstration is generally considered to be used for teaching (McEllin et al., 2017; Saint-Georges et al., 2013; Schaik, Meyer, Ham, \& Hunnius, 2019), specially in music performance, it may not be an effective strategy to perform slowly as changing the tempo might give another interpretation to the music.

Overall, the results indicated that expert pianists seemed to prioritise the teaching of articulation. One possibility is that articulation was more important for the specific piece we selected or the piece itself naturally invited some implementation of articulation. In our previous study, we observed most participants implemented articulation (particularly legato) when they were asked to perform the piece even when neither of articulation nor dynamics was notated on the sheet music. Therefore, it is possible that expert pianists were particularly sensitive to the lack of articulation in the recordings and trying to highlight it. Another possibility is that dynamics modulations in the recordings were too subtle to be noticed by participants and therefore they did not modulate the dynamics aspect of their performance.

One of the limitations of the current study is that participants may not have had a clear idea about what to teach because they did not have a normative ideal performance in mind. As some participants reported that ``(students had) very different interpretations of the melody'' or ``all kids played rhythmically correct'' in the questionnaire, it is plausible that the lack of articulation and/or dynamics might not have been perceived as errors. One solution would be to provide a model performance where both articulation and dynamics are implemented and ask participants (expert pianists) to listen to it first so that they can detect errors by comparing each recording with the model performance.

The reason why participants considered different recording variations as interpretations rather than errors might stem from the fact that all the recordings did not include any pitch errors and were performed the piece with a stable tempo. This was to make sure that only articulation and dynamics features of varied. However, some of the recordings might not have sounded as if unskillful students were playing. In order to create more realistic novice performances, it might be useful to add some jitters to the tempo of the recordings we created, or to ask beginner pianists to play the piece and extract the temporal features of their performances.
